---
title: "Homework #6"
author: "Chanchal Saratkar"
date: "2024-02-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. "Fake data"
```{r}
 library(ggplot2)
 library(MASS)
```

## Generate data

```{r}
# z <- rnorm(n=3000,mean=0.2)
# z <- data.frame(1:3000,z)
# names(z) <- list("ID","myVar")
# z <- z[z$myVar>0,]
# str(z)
# summary(z$myVar)
```
## Histogram
```{r}
# p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
#   geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
# print(p1)
```

## Empirical density curve
```{r}
# p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
# print(p1)
```

## Parameters for `normal` probability density
```{r}
# normPars <- fitdistr(z$myVar,"normal")
# print(normPars)
# str(normPars)
# normPars$estimate["mean"]
```

## `normal` probability density
```{r}
# meanML <- normPars$estimate["mean"]
# sdML <- normPars$estimate["sd"]
# 
# xval <- seq(0,max(z$myVar),len=length(z$myVar))
# 
#  stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
#  p1 + stat
```

## `exponential` probability density
```{r}
# expoPars <- fitdistr(z$myVar,"exponential")
# rateML <- expoPars$estimate["rate"]
# 
# stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
#  p1 + stat + stat2
```

## `uniform` probability density
```{r}
# stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
#  p1 + stat + stat2 + stat3
```

## `gamma` probability density
```{r}
# gammaPars <- fitdistr(z$myVar,"gamma")
# shapeML <- gammaPars$estimate["shape"]
# rateML <- gammaPars$estimate["rate"]
# 
# stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
# p1 + stat + stat2 + stat3 + stat4
```

## `beta` probability density
```{r}
# pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
#   geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
#   xlim(c(0,1)) +
#   geom_density(size=0.75,linetype="dotted")
# 
# betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
# shape1ML <- betaPars$estimate["shape1"]
# shape2ML <- betaPars$estimate["shape2"]
# 
# statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
# pSpecial + statSpecial
```

# 2. My data

## Read data

```{r warning=FALSE}
z <- read.csv("https://raw.githubusercontent.com/ZoopEcoEvo/champlain_seasonality/main/Output/Data/full_data.csv",header=TRUE,sep=",")
str(z)
summary(z)
```

## Library packages
```{r}
library(ggplot2) 
library(MASS)
library(tidyverse)
```


## Histogram
```{r message=FALSE, warning=FALSE}
p1 <- ggplot(data=z, aes(x=size, y = ..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.1) +
  labs(x = "Size (mm)", y = "Density") +
  theme_minimal()
print(p1)
```

## Empirical density curve
```{r message=FALSE, warning=FALSE}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

## Parameters for `normal` probability density
```{r message=FALSE, warning=FALSE}
z <- na.omit(z)
normPars <- fitdistr(z$size,"normal")
print(normPars)
str(normPars)
print(normPars$estimate["mean"])
```

## `normal` probability density
```{r message=FALSE, warning=FALSE}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$size),len=length(z$size))

stat <- stat_function(aes(x = size, y = ..y..), fun = dnorm, colour="red", n = length(z$size), args = list(mean = meanML, sd = sdML))
p1 + stat
```

## `exponential` probability density
```{r message=FALSE, warning=FALSE}
expoPars <- fitdistr(z$size,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = size, y = ..y..), fun = dexp, colour="blue", n = length(z$size), args = list(rate=rateML))
p1 + stat + stat2
```

## `uniform` probability denisty
```{r message=FALSE, warning=FALSE}
stat3 <- stat_function(aes(x = size, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$size), args = list(min=min(z$size), max=max(z$size)))
p1 + stat + stat2 + stat3
```

## `gamma` probability denisty
```{r message=FALSE, warning=FALSE}
gammaPars <- fitdistr(z$size,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = size, y = ..y..), fun = dgamma, colour="brown", n = length(z$size), args = list(shape=shapeML, rate=rateML))
p1 + stat + stat2 + stat3 + stat4
```

## `beta` probability density
```{r message=FALSE, warning=FALSE}
pSpecial <- ggplot(data=z, aes(x=size/(max(size + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.1) + 
  xlim(c(min=min(z$size), max=max(z$size))) +
  geom_density(size=0.75,linetype="dotted") +
  theme_minimal()

betaPars <- fitdistr(x=z$size/max(z$size + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = size, y = ..y..), fun = dbeta, colour="orchid", n = length(z$size), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

# 3. Best fitting distribution

The `normal` and `gamma` probability distributions are the best fitting distributions. There is very little difference between them. `gamma` is slightly better because its peak is closer to the maximum y-values.

# 4. New data set

## `gamma` parameters
```{r message=FALSE, warning=FALSE}
#Finding gamma parameters for the new generated data
gammaPars <- fitdistr(z$size,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

z2 <- rgamma(271, shape = 36.47591, rate = 45.64064)
z2 <- data.frame(1:271, z2)
names(z2) <- list("ID", "size")

# Plotting
p2 <- ggplot(data=z2, aes(x=size, y = ..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.1) +
  labs(x = "Size (mm)", y = "Density") +
  theme_minimal() +
  xlim(.5, 2.25)

stat5 <- stat_function(aes(x = size, y = ..y..), fun = dgamma, colour="brown", n = length(z$size), args = list(shape=shapeML, rate=rateML))

p2 + stat5
```

## Orginal data
```{r message=FALSE, warning=FALSE}
p1 + stat4
```

